---
title: "ANA 515 Assignment 2"
author: "Sravanivennavaram"
date: "06/12/2022"
output:
  html_document:
    df_print: paged
  theme:
    bootswatch: cerulean
    df_print: paged
  word_document: null
  pdf_document: default
---

#### Section 1: Data Description

The Gun Deaths Dataset entails the details of the people who died from gun shots. Such details include the time (year and month) of death, place of death, reason (intent) of death, whether police was involved or was the one whose shot is responsible for the victim's death, gender, race and educational qualification.
This dataset can prove to be highly useful to analyze crime rates as well as suicide rates and take preventive measures by drilling down into the details of the individual involved in the crime or suicide. Technically, if the police is involved, then definitely that instance of gun death is a crime, and hence crime can be analyzed by further drilling down into the personal details of the victim. Similarly, if the intent of gun death is a suicide, the suicide can be analyzed likewise.


#### Section 2: Reading the data into R

```{r, echo = TRUE}
# reading the data using read.csv() function
url <- 'https://raw.githubusercontent.com/fivethirtyeight/guns-data/master/full_data.csv'
gun_deaths <- read.csv(url)
```


#### Section 3: Data Cleaning

```{r, echo = TRUE}
# filtering out the column named, 'X' and 'Hispanic'
gun_deaths <- gun_deaths[, c('year', 'month', 'intent', 'police', 'sex', 'age', 'race', 'place', 'education')]

# filtering out only those instances that involve the police shooting the victim to death (crime)
crime_instances <- gun_deaths[gun_deaths$police == 1, c('year', 'month', 'intent', 'sex', 'age', 'race', 'place', 'education')]
n_rows <- nrow(crime_instances)
n_cols <- ncol(crime_instances)
```


#### Section 4: Characteristics of the data

This dataframe has `r n_rows` and `r n_cols`. The names of the columns and a brief description of each are in the table below:

```{r, echo = TRUE}
data_description <- data.frame(
                               `Column Name` = c('year', 'month', 'intent', 'sex', 'age', 'race', 'place', 'education'),
                                Description =  c('The year when the shooting takes place',
                                                 'The month when the shooting takes place',
                                                 'The motive behind the shooting',
                                                 'Gender of the victim',
                                                 'Age of the victim',
                                                 'Race of the victim',
                                                 'Scene of crime/shooting',
                                                 'Educational Qualification of the victim')
                              )
data_description
```


#### Section 5: Summary Statistics

```{r, echo = TRUE}
# picking 3 columns of the crime dataset
crime_data_filtered <- crime_instances[, c('sex', 'age', 'education')]
summary_statistics <- summary(crime_data_filtered)
summary_statistics
```